2.3 Prob Model for PCR Positivity Estimation

### 2024-02-21 2_scripts/2.3 Prob Model for PCR pos.qmd
CBS processing and figs, APL/Canpath will be added later

## Load packages
```{r}
library(dplyr)         # Data management
library(ggplot2)       # ggplot
library(ggpubr)     #ggplot2 QQ plot for normality check
library(zoo)           # year-month format
library(lmtest)        # for LR test, comparing models
#library(reshape2)     # for melt() function
library(readxl)        # for reading .xlsx data
library(nnet)          # for multinomial log regression
```

@@

# 12 provinces, recatgorized into five regions
CBSs$region<-rep(NA, length(CBSs$pid))
CBSs$region[CBSs$province %in% c("NL", "PE", "NS", "NB")]<-"Atlantic"
CBSs$region[CBSs$province %in% c("QC", "ON")]<-"Central"
CBSs$region[CBSs$province %in% c("AB", "MB", "SK")]<-"Prairie"
CBSs$region[CBSs$province %in% c("BC")]<-"BC"
CBSs$region[CBSs$province %in% c("NU/NT", "YT")]<-"North"

table(CBSs$province, CBSs$region, useNA = 'ifany')

@@

# Creat a ever inf indicator
CBSs$ever_inf <- 1-CBSs$never_inf
```

@@

tb_ce<-CBSs[CBSs$region=="Central",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_ce$region<-rep("Central", length(tb_ce$n))

tb_pr<-CBSs[CBSs$region=="Prairie",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_pr$region<-rep("Prairie", length(tb_pr$n))

tb_no<-CBSs[CBSs$region=="North",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_no$region<-rep("North", length(tb_no$n))

tb_all<-rbind.data.frame(tb_bc, tb_pr, tb_ce, tb_at, tb_no)
```

>>>>

## Load packages
```{r}
library(dplyr)         # Data management
library(ggplot2)       # ggplot2
library(ggpubr)        # ggplot2 QQ plot for normality check
library(zoo)           # year-month format
library(lmtest)        # for LR test, comparing models
#library(reshape2)     # for melt() function
library(readxl)        # for reading .xlsx data
library(nnet)          # for multinomial log regression
```

@@

# 12 provinces, recatgorized into six regions
# region divided according to Yuan's study. In CBS data since the number of samples from QC and North prov are too small (N=13, 85+83). We decided to exclude samples from those two prov from analysis.
# Also, since APL provides AB data only, we will divide prairie prov into AB and SKS+MN. This should apply to Canpath data as well.

CBSs$region<-rep(NA, length(CBSs$pid))
CBSs$region[CBSs$province %in% c("NL", "PE", "NS", "NB")]<-"Atlantic"
CBSs$region[CBSs$province %in% c("ON")]<-"Ontario"
CBSs$region[CBSs$province %in% c("AB")]<-"Alberta"
CBSs$region[CBSs$province %in% c("MB", "SK")]<-"SK & MB"
CBSs$region[CBSs$province %in% c("BC")]<-"BC"

# In this way, QC, NU/NT, and YT will be coded as 'NA'
table(CBSs$province, CBSs$region, useNA = 'ifany')

@@

# Creat a ever inf indicator
CBSs$ever_inf <- 1-CBSs$never_inf
```
#### Save the dataset
```{r}
save(CBSs, file='../1_data/private/CBSs.RData')
```

@@

tb_on<-CBSs[CBSs$region=="Ontario",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_on$region<-rep("Ontario", length(tb_on$n))

tb_ab<-CBSs[CBSs$region=="Alberta",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_ab$region<-rep("Alberta", length(tb_ab$n))

tb_skm<-CBSs[CBSs$region=="SK & MB",] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T))
tb_skm$region<-rep("SK & MB", length(tb_skm$n))

tb_all<-rbind.data.frame(tb_bc, tb_ab, tb_skm, tb_on, tb_at)
# clean NA's
tb_all<-tb_all[!is.na(tb_all$samplemonth),]
```
